Using spatially biased variables in ecosystem condition accounting with a GIS based workflow

### 2024-01-29 manuscript/HIAs.qmd
styler

# Conditional file directory
dir <- substr(getwd(), 1,2)

@@

path_infrastructure <- ifelse(dir == "C:", 
        "R:/GeoSpatialData/Utility_governmentalServices/Norway_Infrastructure_Index/Original/Infrastrukturindeks_UTM33/infra_tiff.tif",
        "/data/R/GeoSpatialData/Utility_governmentalServices/Norway_Infrastructure_Index/Original/Infrastrukturindeks_UTM33/infra_tiff.tif")

path_naturetypes <- ifelse(dir == "C:", 
               "R:/GeoSpatialData/Habitats_biotopes/Norway_Miljodirektoratet_Naturtyper_nin/Original/Naturtyper_nin_0000_norge_25833_FILEGDB/Naturtyper_nin_0000_norge_25833_FILEGDB.gdb",
               "/data/R/GeoSpatialData/Habitats_biotopes/Norway_Miljodirektoratet_Naturtyper_nin/Original/Naturtyper_nin_0000_norge_25833_FILEGDB/Naturtyper_nin_0000_norge_25833_FILEGDB.gdb")

path_muni <- paste0("../data/Basisdata_0000_Norge_25833_Kommuner_FGDB.gdb")

```

@@

#naturetypes_summary2 <- naturetypes_summary %>%
#  rowwise() %>%
#  mutate(keepers = sum(c_across(
#    all_of(myVars))>0, na.rm=T)) %>%
#  filter(keepers >0,
#         Ecosystem == "våtmark") %>%
#  select(Nature_type, NiN_mainType, Year)
#
#naturetypes_summary2$Nature_type

# Getting the nature types that have the ADSV variables
myVars2 <- c('7TK', '7SE', 'PRTK', 'PRSL')
#naturetypes_summary3 <- naturetypes_summary %>%
#  rowwise() %>%
#  mutate(keepers = sum(c_across(
#    all_of(myVars2))>0, na.rm=T)) %>%
#  filter(keepers >0,
#         Ecosystem == "våtmark") %>%
#  select(Nature_type, NiN_mainType, Year)
#
#naturetypes_summary3$Nature_type
```

@@

# These are the nature types I'm keeping for 7FA.
# Note the absence of ombrotriphic types
alienNTS <- c("Slåttemyr",
              "Semi-naturlig våteng",
              "Rik åpen jordvannsmyr i mellomboreal sone",
              "Rik åpen sørlig jordvannsmyr",
              "Rik åpen jordvannsmyr i nordboreal og lavalpin sone")

# And these are the nature types that we use for the ADSV indicator
 advsNTs <- c(# fens
              "Slåttemyr",
              "Sørlig slåttemyr",
              "Rik åpen jordvannsmyr i mellomboreal sone",
              "Rik åpen sørlig jordvannsmyr",
              "Semi-naturlig myr",
              "Rik åpen jordvannsmyr i nordboreal og lavalpin sone",
              #"Sørlig kaldkilde", # excluded springs

              # bogs (mainly)
              "Høgereligende og nordlig nedbørsmyr",
              "Øyblandingsmyr",
              "Konsentrisk høymyr",
              "Sørlig nedbørsmyr",
              "Atlantisk høymyr",
              "Terrengdekkende myr",
              "Eksentrisk høymyr",
              "Kanthøymyr",
              "Platåhøymyr",
              "Palsmyr")

```

@@

naturetypes <- naturetypes |>
  # keep only wetlands and our selected nature types
  filter(hovedøkosystem == "våtmark",
         naturtype %in% c(advsNTs, alienNTS)) |>
  # calculate the areas (m2) of the polygons
  mutate(area = SHAPE |> st_area()) |>
  # the variable codes and values are all in the same column
  separate_rows(ninBeskrivelsesvariable, sep=",") |>
  separate(col = ninBeskrivelsesvariable,
           into = c("NiN_variable_code", "NiN_variable_value"),
           sep = "_",
           remove=F) |>
  mutate(NiN_variable_value = as.numeric(NiN_variable_value)) |>
  filter(NiN_variable_code %in% c(myVars, myVars2)) |>
  select(id = identifikasjon_lokalId,
         municipality = kommunenummer,
         year = kartleggingsår,
         mosaic = mosaikk,
         quality = lokalitetskvalitet,
         biodiversity = naturmangfold,
         condition = tilstand,
         natureType = naturtype,
         variable = NiN_variable_code,
         value = NiN_variable_value,
         area) |>
  st_transform(myCRS) # Chosing this to match the EDM (se further down)

@@

naturetypes |>
  saveRDS("../data/surveyData.RDS")
#head(naturetypes)
```

@@

# find the correct layer
#st_layers(path_muni)

@@

# This returns some multisurfaces as well
#table(st_geometry_type(muni))
# function to make sure that multipolygons are returned 
ensure_multipolygons <- function(X) {
    tmp1 <- tempfile(fileext = ".gpkg")
    tmp2 <- tempfile(fileext = ".gpkg")
    st_write(X, tmp1)
    gdalUtilities::ogr2ogr(tmp1, tmp2, f = "GPKG", nlt = "MULTIPOLYGON")
    Y <- st_read(tmp2)
    st_sf(st_drop_geometry(X), geom = st_geometry(Y))
}

muni <- ensure_multipolygons(muni)
#table(st_geometry_type(muni)) #OK

# subset of the three target municipalities
muni3 <- muni |>
  filter(kommunenummer %in% c("3020", #Nordre Follo
                              "5001", # Trondheim
                              "5418"  # Målselv
                              )) |>
  mutate(Municipality = case_when(
    kommunenummer == "3020" ~ "Nordre Follo",
    kommunenummer == "5001" ~ "Trondheim",
    kommunenummer == "5418" ~ "Målselv"
  ))

@@

# calculate stats - terrestrial area
terrestrial <- terrestrial |>
  mutate(area_t = geometry |> st_area(),
         t_area_km = 
           round(units::drop_units(area_t * 1e-6)))|>
  as_tibble()
```

@@

tmap_mode("plot")
positionMap <- tm_shape(outline)+
  tm_polygons() +
  tm_shape(centroids)+
    tm_text(text = "Municipality",
      size = 1,
      xmod = 2,
      ymod=-1)+
  tm_symbols(size = 2,
             shape = 17,
             col="black")+
  tm_grid(projection = 4326)+
  tm_layout(bg.color = "skyblue")
```

@@

# The cropping is correct:
#tmap_mode("plot")
#tm_shape(mire_test)+
#  tm_raster(palette = "red")+
#  tm_shape(outline)+
#  tm_borders()+
#  tm_shape(mire_test_nf)+
#  tm_raster(col="red")

@@

mire_nf <- read_stars(mireFile, proxy = T) |>
  setNames("mire") |>
  #mutate(mire = na_if(mire, 0)) |>
  st_crop(nf) |>
  st_as_stars()
# The cropped raster is for some reason far into the ocean!

# The cells have values, so it seems like there is something 
# that happens after the cropping
tmap_mode("plot")
tm_shape(mire_test)+
  tm_raster(palette = "red")+
  tm_shape(outline)+
  tm_borders()+
  tm_shape(mire_nf)+
  tm_raster(col="red")
```

@@

# Plot to check overlap
#ggplot()+
#  geom_spatraster(data = mire_nf_terra)+
#  geom_spatvector(data = nf_vect,
#                  fill = NA)
# The cropping and masking worked.

@@

#tm_shape(mire_test)+
#  tm_raster(palette = "red")+
#  tm_shape(outline)+
#  tm_borders()+
#  tm_shape(mire_nf_stars)+
#  tm_raster(col="red")
# OK

ggplot()+
  geom_stars(data = mire_nf_stars)+
  coord_equal()+
  theme_void()+
  scale_fill_viridis_c()
```

@@

# calculate area of survey coverage maps
# values goes into summary table in the ms
dk2 <-   read_sf("../data/dekningskartNF/Naturtyper NiN - Dekningskart.shp") |>
  rbind(read_sf("../data/dekningskartTR/Naturtyper NiN - Dekningskart.shp")) |>
  rbind(read_sf("../data/dekningskartME/Naturtyper NiN - Dekningskart.shp")) |>
  st_transform(st_crs(muni3)) |>
  st_intersection(muni3) |>
  st_cast("POLYGON") |>
  group_by(Municipality) |>
  summarise(geometry = st_union(geometry)) |>
  mutate(dk_area_km = geometry |> st_area(),
         dk_area_km = round(units::drop_units(dk_area_km * 1e-6)))
```

@@

# calculate area of the mires in each municipality
# -- Nordre Follo
mireArea <- mire_nf_terra |>
  global(c("mean", "sum"), na.rm=T) |>
  add_column("Municipality" = "Nordre Follo") |>
  mutate(mirePercent = round(mean*100, 1),
         mire_km2 = sum/1e+4)

# Caululate the area of mire inside the coverage maps
# -- Nordre Follo
mire_nf_terra_dk <- mire_nf_terra |>
  terra::mask(dk2 |> filter(Municipality == "Nordre Follo")) |>
  global("sum", na.rm=T) |>
  mutate(mireInSurvey_km2 = sum/1e+4) |>
  add_column(Municipality = "Nordre Follo")

```

@@

# this data is on a 100x100m grid
infra <- stars::read_stars(path_infrastructure) |>
  # , which is mor then we need - warp it to 1x1km
  st_warp(cellsize = c(1000, 1000), 
                                   crs = st_crs(nf), 
                                   use_gdal = TRUE,
                                   method = "average") |>
  setNames("infrastructureIndex") |>
  st_transform(myCRS) |>
   mutate(infrastructureIndex = case_when(
    infrastructureIndex < 1 ~ 0,
    infrastructureIndex < 6 ~ 1,
    infrastructureIndex < 12 ~ 2,
    infrastructureIndex >= 12 ~ 3)) |>
  # takeing away point in the sea
  st_crop(outline)

@@

infra <-  eaTools::ea_homogeneous_area(infra, 
                                groups = infrastructureIndex)

# path to local cahing folder
path_temp <- ifelse(dir == "C:", 
        "P:/41201785_okologisk_tilstand_2022_2023/data/cache/",
        "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/cache/")

@@

# Turn m2 into km2
# and sum the total area per HIA
(infraMuni3_tbl <- infraMuni3 |>
  as.data.frame() |>
  mutate(area_HIA_km2 = units::drop_units(area)*1e-6) |>
  group_by(Municipality, infrastructureIndex) |>
  summarise(total_area_HIAs_km2 = round(sum(area_HIA_km2))))

# Calculate the area weighted mean HIA value per municipality
infraMuni3_summary <- infraMuni3_tbl |>
  group_by(Municipality) |>
  summarise(meanHIA = 
              round(
                weighted.mean(
                  infrastructureIndex, total_area_HIAs_km2), 2))

# Make a plot to check that it has worked 
infraMuni3_tbl |>
  ggplot()+
  geom_bar(aes(x = infrastructureIndex, 
               y = total_area_HIAs_km2,
               fill = factor(infrastructureIndex),
               colour = factor(infrastructureIndex)),
           stat = "identity",
           lwd=1.2
           )+
  scale_fill_manual(values = RColorBrewer::brewer.pal(4, "YlOrBr"))+
  scale_color_manual(values = RColorBrewer::brewer.pal(5, "YlOrBr")[-1])+
  facet_grid(cols = vars(Municipality))+
  theme_minimal_hgrid()+
  labs(x = "Human Impact Area (HIA) class",
       y = "Area (km<sup>2</sup>)")+
  theme(axis.title.x = element_textbox_simple(
      width = NULL,
      padding = margin(4, 4, 4, 4),
      margin = margin(4, 0, 0, 0),
      linetype = 1,
      r = grid::unit(8, "pt"),
      fill = "azure1"
    ),
    axis.title.y = element_textbox_simple(
      width = NULL,
      padding = margin(4, 4, 4, 4),
      margin = margin(4, 0, 0, 0),
      linetype = 1,
      orientation = "left-rotated",
      r = grid::unit(8, "pt"),
      fill = "azure1"),
    strip.background = element_blank(),
    strip.text = element_textbox(
      size = 12,
      color = "white", fill = "#5D729D", box.color = "#4A618C",
      halign = 0.5, linetype = 1, r = unit(5, "pt"), width = unit(1, "npc"),
      padding = margin(2, 0, 1, 0), margin = margin(3, 3, 3, 3)
    )
    )+
guides(fill= "none", colour = "none")
```

@@

infraMuniMap <- tm_shape(muni3)+
  tm_borders()+
tm_shape(infraMuni3)+
  tm_polygons(col= "infrastructureIndex",
              style="cat",
              title = "Human Impact\nArea Class")+
  tm_shape(muni3)+
  tm_borders(lwd=3, col="black")+
  tm_facets(by = "Municipality")+
  tm_layout(legend.show=TRUE)

# Plot
muniPlot <- tm_shape(muni3)+
    tm_borders()+
  tm_facets(by = "Municipality",
            ncol=3)+
  tm_shape(terrestrial)+
    tm_fill(col = "lightgreen",
             alpha=.4)+
  tm_shape(ocean)+
    tm_polygons(col = "skyblue",
                border.col = "black")+
  tm_shape(dk)+
    tm_polygons(col = "grey",
                alpha=.8)+
  tm_shape(nature3)+
  tm_polygons(col = "red",
              border.col = "red",
              lwd=3)

(methodsMap <- tmap_arrange(
                          #positionMap, 
                           muniPlot,
                          infraMuniMap,
             widths = c(.4, .6)))

#cowplot::plot_grid(tmap_grob(muniPlot),
#                   tmap_grob(infraMuniMap),
#                   ncol=1,)

@@

# Take muni3 and add all sorts of other data to it
# using left_join.
# Save local cached file
muni_tbl <- muni3 |>
  # calculate total area
  mutate(area_km = 
           round(
             units::drop_units(
               geom |> st_area() * 1e-6))) |>
  # make tibble for the coming join
  as_tibble() |>
  # paste inn terrestrial area
  left_join(terrestrial |> select(kommunenummer, t_area_km),
            keep=F) |>
  # area of survey
  left_join(dk2 |> select(Municipality, dk_area_km)) |>
  mutate(dk_percent = round((dk_area_km/t_area_km) * 100)) |>
  # number of polygons
  left_join(nature3_tbl |> select(Municipality, n)) |>
  # total mire area and %
  left_join(mireArea |> select(Municipality, mirePercent, mire_km2)) |>
  # % mire inside survey coverage map
  left_join(mire_nf_terra_dk |> select(Municipality, mireInSurvey_km2)) |>
  mutate(mireInSurvery_percent = round(mireInSurvey_km2/mire_km2*100, 2)) |>
  left_join(infraMuni3_summary)

@@

| Variable code        | Variable name                                    | Measurement unit                                                                        | Description                                                                                                    | Reference                          |
|---------------|---------------|---------------|---------------|---------------|
| 7FK                  | Prevalence of alien species                      | Unitless, ordinal, non-linear scale from 1 (no alien species) to 7 (only alien species) | The fraction of the species composition made up from alien species                                             | @runehalvorsen2019                 |
| 7SE                  | Human caused abration or abration-caused erosion | Unitless, ordinal, non-linear scale from 1 to 4.                                        | Measures the frequency of imagined 4 m^2^ quadrats layed over the area that has some sign of abration          | @runehalvorsen2019                 |
| PRSL                 | *as above*                                       | Unitless, ordinal, non-linear scale from 0 to 7.                                        | Same as 7SE, but recorded at a higher resolution                                                               | @miljødirektoratet                 |
| 7TK                  | Tracks from large vehicles                       | Unitless, ordinal, non-linear scale from 1 to 4.                                        | Measures the frequency of imagined 100 m^2^ quadrats layed over the area that has some signs of vehicle tracks | @rune_halvorsen_dokumentasjon_2019 |
| PRTK                 | *as above*                                       | Unitless, ordinal, non-linear scale from 0 to 7.                                        | Same as 7TK, but recorded at a higher resolution                                                               | @miljødirektoratet                 |
| Infrastructure Index | Infrastructure Index                             | Unitelss linear scale from 0 to 13.2                                                    | Unitelss index ranging from from 0 to 13.2                                                                     | @erikstad_index_2023               |
| HIA                  | Human Impact area                                | Ordinal, non-linear scale from 1 to 4                                                   | A categorical representation of the infrastructure index                                                       | this paper                         |

>>>>

# Conditional file directory
dir <- substr(getwd(), 1, 2)

@@

path_infrastructure <- ifelse(dir == "C:",
  "R:/GeoSpatialData/Utility_governmentalServices/Norway_Infrastructure_Index/Original/Infrastrukturindeks_UTM33/infra_tiff.tif",
  "/data/R/GeoSpatialData/Utility_governmentalServices/Norway_Infrastructure_Index/Original/Infrastrukturindeks_UTM33/infra_tiff.tif"
)

path_naturetypes <- ifelse(dir == "C:",
  "R:/GeoSpatialData/Habitats_biotopes/Norway_Miljodirektoratet_Naturtyper_nin/Original/Naturtyper_nin_0000_norge_25833_FILEGDB/Naturtyper_nin_0000_norge_25833_FILEGDB.gdb",
  "/data/R/GeoSpatialData/Habitats_biotopes/Norway_Miljodirektoratet_Naturtyper_nin/Original/Naturtyper_nin_0000_norge_25833_FILEGDB/Naturtyper_nin_0000_norge_25833_FILEGDB.gdb"
)

path_muni <- paste0("../data/Basisdata_0000_Norge_25833_Kommuner_FGDB.gdb")
```

@@

# naturetypes_summary2 <- naturetypes_summary %>%
#  rowwise() %>%
#  mutate(keepers = sum(c_across(
#    all_of(myVars))>0, na.rm=T)) %>%
#  filter(keepers >0,
#         Ecosystem == "våtmark") %>%
#  select(Nature_type, NiN_mainType, Year)
#
# naturetypes_summary2$Nature_type

# Getting the nature types that have the ADSV variables
myVars2 <- c("7TK", "7SE", "PRTK", "PRSL")
# naturetypes_summary3 <- naturetypes_summary %>%
#  rowwise() %>%
#  mutate(keepers = sum(c_across(
#    all_of(myVars2))>0, na.rm=T)) %>%
#  filter(keepers >0,
#         Ecosystem == "våtmark") %>%
#  select(Nature_type, NiN_mainType, Year)
#
# naturetypes_summary3$Nature_type
```

@@

# These are the nature types I'm keeping for 7FA.
# Note the absence of ombrotriphic types
alienNTS <- c(
  "Slåttemyr",
  "Semi-naturlig våteng",
  "Rik åpen jordvannsmyr i mellomboreal sone",
  "Rik åpen sørlig jordvannsmyr",
  "Rik åpen jordvannsmyr i nordboreal og lavalpin sone"
)

# And these are the nature types that we use for the ADSV indicator
advsNTs <- c( # fens
  "Slåttemyr",
  "Sørlig slåttemyr",
  "Rik åpen jordvannsmyr i mellomboreal sone",
  "Rik åpen sørlig jordvannsmyr",
  "Semi-naturlig myr",
  "Rik åpen jordvannsmyr i nordboreal og lavalpin sone",
  # "Sørlig kaldkilde", # excluded springs

  # bogs (mainly)
  "Høgereligende og nordlig nedbørsmyr",
  "Øyblandingsmyr",
  "Konsentrisk høymyr",
  "Sørlig nedbørsmyr",
  "Atlantisk høymyr",
  "Terrengdekkende myr",
  "Eksentrisk høymyr",
  "Kanthøymyr",
  "Platåhøymyr",
  "Palsmyr"
)
```

@@

naturetypes <- naturetypes |>
  # keep only wetlands and our selected nature types
  filter(
    hovedøkosystem == "våtmark",
    naturtype %in% c(advsNTs, alienNTS)
  ) |>
  # calculate the areas (m2) of the polygons
  mutate(area = SHAPE |> st_area()) |>
  # the variable codes and values are all in the same column
  separate_rows(ninBeskrivelsesvariable, sep = ",") |>
  separate(
    col = ninBeskrivelsesvariable,
    into = c("NiN_variable_code", "NiN_variable_value"),
    sep = "_",
    remove = F
  ) |>
  mutate(NiN_variable_value = as.numeric(NiN_variable_value)) |>
  filter(NiN_variable_code %in% c(myVars, myVars2)) |>
  select(
    id = identifikasjon_lokalId,
    municipality = kommunenummer,
    year = kartleggingsår,
    mosaic = mosaikk,
    quality = lokalitetskvalitet,
    biodiversity = naturmangfold,
    condition = tilstand,
    natureType = naturtype,
    variable = NiN_variable_code,
    value = NiN_variable_value,
    area
  ) |>
  st_transform(myCRS) # Chosing this to match the EDM (se further down)

@@

naturetypes |>
  saveRDS("../data/surveyData.RDS")
# head(naturetypes)
```

@@

# find the correct layer
# st_layers(path_muni)

@@

# This returns some multisurfaces as well
# table(st_geometry_type(muni))
# function to make sure that multipolygons are returned
ensure_multipolygons <- function(X) {
  tmp1 <- tempfile(fileext = ".gpkg")
  tmp2 <- tempfile(fileext = ".gpkg")
  st_write(X, tmp1)
  gdalUtilities::ogr2ogr(tmp1, tmp2, f = "GPKG", nlt = "MULTIPOLYGON")
  Y <- st_read(tmp2)
  st_sf(st_drop_geometry(X), geom = st_geometry(Y))
}

muni <- ensure_multipolygons(muni)
# table(st_geometry_type(muni)) #OK

# subset of the three target municipalities
muni3 <- muni |>
  filter(kommunenummer %in% c(
    "3020", # Nordre Follo
    "5001", # Trondheim
    "5418" # Målselv
  )) |>
  mutate(Municipality = case_when(
    kommunenummer == "3020" ~ "Nordre Follo",
    kommunenummer == "5001" ~ "Trondheim",
    kommunenummer == "5418" ~ "Målselv"
  ))

@@

# calculate stats - terrestrial area
terrestrial <- terrestrial |>
  mutate(
    area_t = geometry |> st_area(),
    t_area_km =
      round(units::drop_units(area_t * 1e-6))
  ) |>
  as_tibble()
```

@@

tmap_mode("plot")
positionMap <- tm_shape(outline) +
  tm_polygons() +
  tm_shape(centroids) +
  tm_text(
    text = "Municipality",
    size = 1,
    xmod = 2,
    ymod = -1
  ) +
  tm_symbols(
    size = 2,
    shape = 17,
    col = "black"
  ) +
  tm_grid(projection = 4326) +
  tm_layout(bg.color = "skyblue")
```

@@

# The cropping is correct:
# tmap_mode("plot")
# tm_shape(mire_test)+
#  tm_raster(palette = "red")+
#  tm_shape(outline)+
#  tm_borders()+
#  tm_shape(mire_test_nf)+
#  tm_raster(col="red")

@@

mire_nf <- read_stars(mireFile, proxy = T) |>
  setNames("mire") |>
  # mutate(mire = na_if(mire, 0)) |>
  st_crop(nf) |>
  st_as_stars()
# The cropped raster is for some reason far into the ocean!

# The cells have values, so it seems like there is something
# that happens after the cropping
tmap_mode("plot")
tm_shape(mire_test) +
  tm_raster(palette = "red") +
  tm_shape(outline) +
  tm_borders() +
  tm_shape(mire_nf) +
  tm_raster(col = "red")
```

@@

# Plot to check overlap
# ggplot()+
#  geom_spatraster(data = mire_nf_terra)+
#  geom_spatvector(data = nf_vect,
#                  fill = NA)
# The cropping and masking worked.

@@

# tm_shape(mire_test)+
#  tm_raster(palette = "red")+
#  tm_shape(outline)+
#  tm_borders()+
#  tm_shape(mire_nf_stars)+
#  tm_raster(col="red")
# OK

ggplot() +
  geom_stars(data = mire_nf_stars) +
  coord_equal() +
  theme_void() +
  scale_fill_viridis_c()
```

@@

# calculate area of survey coverage maps
# values goes into summary table in the ms
dk2 <- read_sf("../data/dekningskartNF/Naturtyper NiN - Dekningskart.shp") |>
  rbind(read_sf("../data/dekningskartTR/Naturtyper NiN - Dekningskart.shp")) |>
  rbind(read_sf("../data/dekningskartME/Naturtyper NiN - Dekningskart.shp")) |>
  st_transform(st_crs(muni3)) |>
  st_intersection(muni3) |>
  st_cast("POLYGON") |>
  group_by(Municipality) |>
  summarise(geometry = st_union(geometry)) |>
  mutate(
    dk_area_km = geometry |> st_area(),
    dk_area_km = round(units::drop_units(dk_area_km * 1e-6))
  )
```

@@

# calculate area of the mires in each municipality
# -- Nordre Follo
mireArea <- mire_nf_terra |>
  global(c("mean", "sum"), na.rm = T) |>
  add_column("Municipality" = "Nordre Follo") |>
  mutate(
    mirePercent = round(mean * 100, 1),
    mire_km2 = sum / 1e+4
  )

# Caululate the area of mire inside the coverage maps
# -- Nordre Follo
mire_nf_terra_dk <- mire_nf_terra |>
  terra::mask(dk2 |> filter(Municipality == "Nordre Follo")) |>
  global("sum", na.rm = T) |>
  mutate(mireInSurvey_km2 = sum / 1e+4) |>
  add_column(Municipality = "Nordre Follo")
```

@@

# this data is on a 100x100m grid
infra <- stars::read_stars(path_infrastructure) |>
  # , which is mor then we need - warp it to 1x1km
  st_warp(
    cellsize = c(1000, 1000),
    crs = st_crs(nf),
    use_gdal = TRUE,
    method = "average"
  ) |>
  setNames("infrastructureIndex") |>
  st_transform(myCRS) |>
  mutate(infrastructureIndex = case_when(
    infrastructureIndex < 1 ~ 0,
    infrastructureIndex < 6 ~ 1,
    infrastructureIndex < 12 ~ 2,
    infrastructureIndex >= 12 ~ 3
  )) |>
  # takeing away point in the sea
  st_crop(outline)

@@

infra <- eaTools::ea_homogeneous_area(infra,
  groups = infrastructureIndex
)

# path to local cahing folder
path_temp <- ifelse(dir == "C:",
  "P:/41201785_okologisk_tilstand_2022_2023/data/cache/",
  "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/cache/"
)

@@

# Turn m2 into km2
# and sum the total area per HIA
(infraMuni3_tbl <- infraMuni3 |>
  as.data.frame() |>
  mutate(area_HIA_km2 = units::drop_units(area) * 1e-6) |>
  group_by(Municipality, infrastructureIndex) |>
  summarise(total_area_HIAs_km2 = round(sum(area_HIA_km2))))

# Calculate the area weighted mean HIA value per municipality
infraMuni3_summary <- infraMuni3_tbl |>
  group_by(Municipality) |>
  summarise(
    meanHIA =
      round(
        weighted.mean(
          infrastructureIndex, total_area_HIAs_km2
        ), 2
      )
  )

# Make a plot to check that it has worked
infraMuni3_tbl |>
  ggplot() +
  geom_bar(
    aes(
      x = infrastructureIndex,
      y = total_area_HIAs_km2,
      fill = factor(infrastructureIndex),
      colour = factor(infrastructureIndex)
    ),
    stat = "identity",
    lwd = 1.2
  ) +
  scale_fill_manual(values = RColorBrewer::brewer.pal(4, "YlOrBr")) +
  scale_color_manual(values = RColorBrewer::brewer.pal(5, "YlOrBr")[-1]) +
  facet_grid(cols = vars(Municipality)) +
  theme_minimal_hgrid() +
  labs(
    x = "Human Impact Area (HIA) class",
    y = "Area (km<sup>2</sup>)"
  ) +
  theme(
    axis.title.x = element_textbox_simple(
      width = NULL,
      padding = margin(4, 4, 4, 4),
      margin = margin(4, 0, 0, 0),
      linetype = 1,
      r = grid::unit(8, "pt"),
      fill = "azure1"
    ),
    axis.title.y = element_textbox_simple(
      width = NULL,
      padding = margin(4, 4, 4, 4),
      margin = margin(4, 0, 0, 0),
      linetype = 1,
      orientation = "left-rotated",
      r = grid::unit(8, "pt"),
      fill = "azure1"
    ),
    strip.background = element_blank(),
    strip.text = element_textbox(
      size = 12,
      color = "white", fill = "#5D729D", box.color = "#4A618C",
      halign = 0.5, linetype = 1, r = unit(5, "pt"), width = unit(1, "npc"),
      padding = margin(2, 0, 1, 0), margin = margin(3, 3, 3, 3)
    )
  ) +
  guides(fill = "none", colour = "none")
```

@@

infraMuniMap <- tm_shape(muni3) +
  tm_borders() +
  tm_shape(infraMuni3) +
  tm_polygons(
    col = "infrastructureIndex",
    style = "cat",
    title = "Human Impact\nArea Class"
  ) +
  tm_shape(muni3) +
  tm_borders(lwd = 3, col = "black") +
  tm_facets(by = "Municipality") +
  tm_layout(legend.show = TRUE)

# Plot
muniPlot <- tm_shape(muni3) +
  tm_borders() +
  tm_facets(
    by = "Municipality",
    ncol = 3
  ) +
  tm_shape(terrestrial) +
  tm_fill(
    col = "lightgreen",
    alpha = .4
  ) +
  tm_shape(ocean) +
  tm_polygons(
    col = "skyblue",
    border.col = "black"
  ) +
  tm_shape(dk) +
  tm_polygons(
    col = "grey",
    alpha = .8
  ) +
  tm_shape(nature3) +
  tm_polygons(
    col = "red",
    border.col = "red",
    lwd = 3
  )

(methodsMap <- tmap_arrange(
  # positionMap,
  muniPlot,
  infraMuniMap,
  widths = c(.4, .6)
))

# cowplot::plot_grid(tmap_grob(muniPlot),
#                   tmap_grob(infraMuniMap),
#                   ncol=1,)

@@

# Take muni3 and add all sorts of other data to it
# using left_join.
# Save local cached file
muni_tbl <- muni3 |>
  # calculate total area
  mutate(
    area_km =
      round(
        units::drop_units(
          geom |> st_area() * 1e-6
        )
      )
  ) |>
  # make tibble for the coming join
  as_tibble() |>
  # paste inn terrestrial area
  left_join(terrestrial |> select(kommunenummer, t_area_km),
    keep = F
  ) |>
  # area of survey
  left_join(dk2 |> select(Municipality, dk_area_km)) |>
  mutate(dk_percent = round((dk_area_km / t_area_km) * 100)) |>
  # number of polygons
  left_join(nature3_tbl |> select(Municipality, n)) |>
  # total mire area and %
  left_join(mireArea |> select(Municipality, mirePercent, mire_km2)) |>
  # % mire inside survey coverage map
  left_join(mire_nf_terra_dk |> select(Municipality, mireInSurvey_km2)) |>
  mutate(mireInSurvery_percent = round(mireInSurvey_km2 / mire_km2 * 100, 2)) |>
  left_join(infraMuni3_summary)

@@

| Variable code        | Variable name                                    | Measurement unit                                                                        | Description                                                                                                    | Reference                          |
|-------------|-------------|---------------|--------------------|-------------|
| 7FK                  | Prevalence of alien species                      | Unitless, ordinal, non-linear scale from 1 (no alien species) to 7 (only alien species) | The fraction of the species composition made up from alien species                                             | @runehalvorsen2019                 |
| 7SE                  | Human caused abration or abration-caused erosion | Unitless, ordinal, non-linear scale from 1 to 4.                                        | Measures the frequency of imagined 4 m^2^ quadrats layed over the area that has some sign of abration          | @runehalvorsen2019                 |
| PRSL                 | *as above*                                       | Unitless, ordinal, non-linear scale from 0 to 7.                                        | Same as 7SE, but recorded at a higher resolution                                                               | @miljødirektoratet                 |
| 7TK                  | Tracks from large vehicles                       | Unitless, ordinal, non-linear scale from 1 to 4.                                        | Measures the frequency of imagined 100 m^2^ quadrats layed over the area that has some signs of vehicle tracks | @rune_halvorsen_dokumentasjon_2019 |
| PRTK                 | *as above*                                       | Unitless, ordinal, non-linear scale from 0 to 7.                                        | Same as 7TK, but recorded at a higher resolution                                                               | @miljødirektoratet                 |
| Infrastructure Index | Infrastructure Index                             | Unitelss linear scale from 0 to 13.2                                                    | Unitelss index ranging from from 0 to 13.2                                                                     | @erikstad_index_2023               |
| HIA                  | Human Impact area                                | Ordinal, non-linear scale from 1 to 4                                                   | A categorical representation of the infrastructure index                                                       | this paper                         |
